---
title: "Visions of a Better WRTA"
author: "Garren Kalter and Luke Watson Sharer"
date: "12/4/2020"
output:
  ioslides_presentation: default
  slidy_presentation: default
---


```{r}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

The WRTA route and stop configuration is selected in a manual manner

The purpose of this project is to prove that geospatial analysis using R
can provide unique insights into the functioning of transit systems.


## Data

The vast majority of our data is derived from WRTA reports, TIGER shapefiles, and 
census data. Some examples can be seen below.

```{r}

```


## Methods

Our methodology is fairly straight forward. For this project we have four 
deliverables. The methodology deployed is as follows:

- Population Density Map (Read in data --> Rasterize it --> Display it)
- Traffic Count Map (Read in --> Rasterize it --> Display it)
- Categorization of Ridership Map (Preprocess --> Read In --> Join with Line Data --> Rasterize --> Create categories --> Display it)
- Routes above maximum subsidy (Preprocess --> Read In --> Join with Line Data --> Rasterize --> Create Categories --> Display it)

## Results

We found that there are many routes which are highly inefficient in Worcester as displayed below:

```{r}

```

## Resullts Cont.
By assessing traffic count, where ridership is already high, and where there are 
dense populations in Worcester one can come to make recommendations.


```{r}

```

## Conclusion

By looking at the above data we recommend cutting the ___ lines and adding those resources to the ____lines. In doing so, the system better serves the Worcester populace. 

For example, leaning on analysis completed for another course,  upwards of 48,000 residents would benefit from more service on Route 27 with 13,000 of those resident fitting the profile of a typical WRTA user.

Further analysis and modeling using packages like zoo lpsolve will be completed in the coming months. We look forward to sharing these findings as they emerge. 
and 



