---
title: "Garren"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Garren}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
#Reading in Data
```{r, eval = FALSE}
ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% 
  readxl::read_excel() 
county <- system.file("extdata/tl_2017_25027_faces/tl_2017_25027_faces.shp",
                     package = "wrta") %>% st_read()
roads <- system.file("extdata/tl_2018_25027_roads/tl_2018_25027_roads.shp",
                     package = "wrta") %>% st_read()
stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()
routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% st_read() %>% st_zm()
```

# Join
```{r, eval = FALSE}
#Identifying whether variables Match
routes$Bus_Route %in% ridership$ROUTE

#completing a join via coercion
routes_ridership <- right_join(x = ridership, 
                               y = routes, by = c("ROUTE" = "BusRoute")) 
```

## Categories Map
```{r, eval = FALSE}
#cleaning data up
distinct_routes <- routes_ridership %>% distinct(ROUTE, .keep_all = TRUE) 

#generating a new field to categorize ridership
distinct_routes_filter$ride_cat
distinct_routes_filter <- distinct_routes %>% 
  mutate(ride_cat = ifelse(WKDAY_RIDERS > 200000, "high", "other"), 
         ride_cat = ifelse(between(WKDAY_RIDERS, 100000, 199999), "medium", ride_cat),
         ride_cat = ifelse(WKDAY_RIDERS < 100000, "low", ride_cat)) %>% 
  st_as_sf()

ggplot(distinct_routes_filter) + geom_sf(aes(color = ride_cat))
```

## Subsidies Map
```{r}

subsidy_routes_filter$sub_cat
subsidy_routes_filter <- distinct_routes %>% 
  mutate(sub_cat = ifelse(subsidy > 10, "high", "other"), 
         sub_cat = ifelse(between(subsidy, 3, 10), "medium", sub_cat),
         sub_cat = ifelse(subsidy < 3, "low", sub_cat)) %>% 
  st_as_sf()

ggplot(subsidy_routes_filter) + geom_sf(aes(color = sub_cat))
```


## Headways Map

```{r}
headways_routes_filter$head_cat
headways_routes_filter <- distinct_routes %>% 
  mutate(head_cat = ifelse(subsidy > 60, "high", "other"), 
         head_cat = ifelse(between(subsidy, 30, 60), "medium", head_cat),
         head_cat = ifelse(subsidy < 30, "low", head_cat)) %>% 
  st_as_sf()

ggplot(headways_routes_filter) + geom_sf(aes(color = head_cat))


```


```{r, eval = FALSE}
#Libraries Used
```{r setup}
library(wrta)
library(sf)
library(dplyr)
library(rgdal)
library(ggplot2)
```

#Reading in Data
```{r, eval = FALSE}
#ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% st_read

ridership <- system.file("extdata/Weekday_ridership.csv", package = "wrta") %>% 
  readr::read_csv(.)

county <- system.file("extdata/Worcester/worcester.shp",
                     package = "wrta") %>% st_read() 

roads <- system.file("extdata/MassDOT_Roads_SHP/EOTMAJROADS_ARC.shp",
                     package = "wrta") %>% st_read() 

stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()

routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% read_sf() %>% st_zm() 


```

#Cropping
```{r, eval = FALSE}
#roads_worc <- st_intersection(x = roads, y = county) 

#stops <- st_transform
#(x = stops, crs = county)
#stops_worc <- st_intersection(x = stops, y = county, crs)

```

# Join
```{r, eval = FALSE}
#Identifying whether variables Match
routes$Bus_Route %in% ridership$ROUTE

routes_spatial <- signature(obj = SpatialLinesDataFrame)

#clean data
routes_c <- routes %>% dplyr::select(BusRoute, Shape_Leng, geometry)

ridershipd <- as.numeric(ridership)

#completing a join via coercion
route_ridership <- left_join(x = ridership, y = routes_c, 
                             by = c("ROUTE" = "BusRoute")) 
plot(route_ridership)
route_ridership_spatial <- SpatialLinesDataFrame(LinesList = "routes", 
                                                 data = "ridership")

```

#Generating New Fields
```{r, eval = FALSE}
#cleaning data up
cat <- ridership %>% distinct(ROUTE, .keep_all = TRUE) 
route_ridership <- 

#generating a new field to categorize ridership
ridership <- ridership %>% 
  mutate(ride_cat = ifelse(WKDAY_RIDERS > 200000, "high", "other"), 
         ride_cat = ifelse(between(WKDAY_RIDERS, 100000, 199999), "medium", ride_cat),
         ride_cat = ifelse(WKDAY_RIDERS < 100000, "low", ride_cat)) %>% st_as_sf
#generating a new field to categorize 
ridership <- ridership %>% 
  mutate(sub_cat = ifelse(subsidy > 10, "high", "other"), 
         sub_cat = ifelse(between(subsidy, 3, 10), "medium", sub_cat),
         sub_cat = ifelse(subsidy < 3, "low", sub_cat)) 

ridership <- ridership %>% as_tibble() %>% 
  full_join(routes_c, by = c("ROUTE" = "BusRoute"))

ridership <- ridership %>% SpatialLinesDataFrame(data = ridership, lines = routes_c)
ridership <- cbind.data.frame(ridership, routes_c)

%>% 
  ggplot() + geom_sf(aes(fill = geometry)
                     
plot(county %>% st_geometry, col ="grey")
plot(routes_ridership %>% st_geometry, col = "green", add = TRUE )

test <- as_Spatial(cat2)
%>% 
  st_geometry()
cat3 <- st_as_sf(cat2, sf_column_name = geometry)

ggplot(cat2) + geom_sf(aes(fill = "ride_cat"))
plot(cat2)


ridership %>% summarize(mean(WKDAY_RIDERS), sd(WKDAY_RIDERS)) #%>% plot()

# mean yields by yield category
districts_mu <- districts2 %>% 
  mutate(yld_cat = ifelse(yield > 3.5, "high", "other"), 
         yld_cat = ifelse(between(yield, 3.5, 4.5), "medium", yld_cat),
         yld_cat = ifelse(yield < 3.5, "low", yld_cat)) %>% 
  dplyr::select(1:5, 7) %>% 
  group_by(yld_cat) %>% summarize(yield = mean(yield))

ggplot(districts_mu) + geom_sf(aes(fill = yield))

# category map
districts2 %>% 
  mutate(yld_cat = ifelse(yield > 3.5, "high", "other"), 
         yld_cat = ifelse(between(yield, 3.5, 4.5), "medium", yld_cat),
         yld_cat = ifelse(yield < 3.5, "low", yld_cat)) %>% 
  ggplot() + geom_sf(aes(fill = yld_cat))

```

# Map Categorizing Routes
```{r, eval = FALSE}
categories <- quantile(raintot, probs = seq(0, 1, 1 / 3)) quan
 # category map
plot(county)
ridership %>% 
  mutate(WKDAY_RIDERS_cat = ifelse(WKDAY_RIDERS > 100000, "high", WKDAY_RIDERS_cat),  
          WKDAY_RIDERS_cat = ifelse(between(WKDAY_RIDERS, 10000, 99999), "medium", WKDAY_RIDERS_cat),
        WKDAY_RIDERS_cat = ifelse(WKDAY_RIDERS < 10000, "low", WKDAY_RIDERS_cat),
        WKDAY_RIDERS_cat = ifelse(WKDAY_RIDERS, "No Route", WKDAY_RIDERS_cat)) %>% 
    ggplot() + geom_sf(aes(fill = yld_cat))

```

# Map Showcasing Routes above Maximum Allowable Subsidy

```{r, eval = FALSE}
> districts2 %>% 
+   mutate(yld_cat = ifelse(yield > , "high", "other") , 
+          yld_cat = ifelse(between(yield, 3.5, 4.5), "medium", yld_cat),
+          yld_cat = ifelse(yield < 3.5, "low", yld_cat)) %>% 
+   ggplot() + geom_sf(aes(fill = yld_cat))
```

