---
title: "WRTA Route Characteristics and Exploratory Findings"
author: "Garren Kalter"
date: "December 6th, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Garren}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following is a vignette of Garren's work for the final project. As you can see below, the following is predominantly vector data that has been cut to Worcester's city boundaries, joined with relevant preprocessed data about the routes, and then displayed to highlight charateristic discrepancies between them. The last two maps indicate on aggregate which routes should be considered for additional funding and which should be considered for closure.

* Many outputs have been shortened using head() in order to improve readability while
still allowing the end user to see the outputs for themselves. A few have been left out
for redundancies sake and/or where doing so didn't make sense.


## Reading in Relevant Libraries

```{r}
library(wrta)
library(sf)
library(dplyr)
library(rgdal)
library(geospaar)
```


## Reading in Relevant Data

```{r, warning = FALSE, message = FALSE }
ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% 
  readxl::read_excel() 
county <- system.file("extdata/tl_2017_25027_faces/tl_2017_25027_faces.shp",
                     package = "wrta") %>% st_read()
roads <- system.file("extdata/tl_2018_25027_roads/tl_2018_25027_roads.shp",
                     package = "wrta") %>% st_read()
stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()
routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% st_read() %>% st_zm()
worcester <- system.file("extdata/Worcester/worcester.shp",
                     package = "wrta") %>% st_read()
```

## Projecting and Cropping to Worcester, MA

```{r, warning = FALSE, message = FALSE}
#reproject
routes <- st_transform(x = routes, crs = st_crs(worcester))
head(routes)

stops <- st_transform(x = stops,crs = st_crs(worcester))
head(stops)

roads <- st_transform(x = roads, crs = st_crs(worcester))
head(roads)

#major roads in Worcester
roads_worc <- st_intersection(x = roads, y = worcester)
head(roads_worc)
#bus stops
stops_worc <- st_intersection(x = stops, y = worcester)
head(stops_worc)
#bus routes
routes_worc <- st_intersection(x = routes, y = worcester)
head(routes_worc)
```

## Joining Preprocessed Data to Routes Shapefile

```{r, warning = FALSE, message = FALSE}
#Identifying whether variables Match
routes$Bus_Route %in% ridership$ROUTE

#completing a join via coercion
routes_ridership <- right_join(x = ridership, 
                               y = routes_worc, by = c("ROUTE" = "BusRoute")) 
head(routes_ridership)
```

## Categories Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
#cleaning data up

distinct_routes <- routes_ridership %>% distinct(ROUTE, .keep_all = TRUE) 
head(distinct_routes)

#generating a new field to categorize ridership
distinct_routes_filter$ride_cat
distinct_routes_filter <- distinct_routes %>% 
  mutate(ride_cat = ifelse(WKDAY_RIDERS > 170000, "high", "other"), 
         ride_cat = ifelse(between(WKDAY_RIDERS, 10000, 170000), "medium", ride_cat),
         ride_cat = ifelse(WKDAY_RIDERS < 10000, "low", ride_cat)) %>% 
  st_as_sf()

head(distinct_routes_filter)
```

### Correcting for Overlaid Routes

```{r, warning = FALSE, message = FALSE}
# uh oh it isn't displaying properly, let's break that up so routes other 
##[2]than the one overlaid on top show.

distinct_route_filter_low <- subset(distinct_routes_filter, 
                                    ride_cat == "low", select = c("geometry"))
distinct_route_filter_low

distinct_route_filter_medium <- subset(distinct_routes_filter, 
                                     ride_cat == "medium", select = c("geometry"))
distinct_route_filter_medium

distinct_route_filter_high <- subset(distinct_routes_filter, 
                                     ride_cat == "high", select = c("geometry"))
distinct_route_filter_high
```

### Plotting Categories of Ridership

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(distinct_route_filter_low %>% st_geometry(), col = "red", add = TRUE)
plot(distinct_route_filter_medium %>% st_geometry(), col = "yellow", add = TRUE)
plot(distinct_route_filter_high %>% st_geometry(), col = "green", add = TRUE)
title(main = "Routes Categorized by Ridership")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("red", "yellow2", "green"), bty = "n")
```

## Subsidies Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
#generating a new field categorizing subsidy

subsidy_routes_filter$sub_cat
subsidy_routes_filter <- distinct_routes %>% 
  mutate(sub_cat = ifelse(subsidy > 10.00, "high", "other"), 
         sub_cat = ifelse(between(subsidy, 3.00, 10.00), "medium", sub_cat),
         sub_cat = ifelse(subsidy < 3.00, "low", sub_cat)) %>% 
  st_as_sf()

head(subsidy_routes_filter)
```

### Correcting for Overlaid Routes

```{r, warning = FALSE, message = FALSE}
# uh oh it isn't displaying properly, let's break that up so routes other 
##[2]than the one overlaid on top show.

subsidy_route_filter_low <- subset(subsidy_routes_filter, 
                                   sub_cat == "low", select = c("geometry"))
subsidy_route_filter_low

subsidy_route_filter_medium <- subset(subsidy_routes_filter, 
                                      sub_cat == "medium", select = c("geometry"))
subsidy_route_filter_medium

subsidy_route_filter_high <- subset(subsidy_routes_filter, 
                                    sub_cat == "high", select = c("geometry"))
subsidy_route_filter_high
```

### Plotting Subsidies Map

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(subsidy_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(subsidy_route_filter_medium %>% st_geometry(), col = "yellow2", add = TRUE)
plot(subsidy_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
title(main = "Routes Categorized by Subsidy")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "yellow2", "red"), bty = "n")
```


## Headways Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
#generating a new field categories headway times. This is how often the bus goes
#[2] by the same stop.

headways_routes_filter$head_cat
headways_routes_filter <- distinct_routes %>% 
  mutate(head_cat = ifelse(time > 60, "high", "other"), 
         head_cat = ifelse(between(time, 30, 60), "medium", head_cat),
         head_cat = ifelse(time < 30, "low", head_cat)) %>% 
  st_as_sf()

head(headways_routes_filter)
```

### Correcting for Overlaid Routes

As you can see below there are no routes which run below 30 minutes, this is a problem when trying to run an efficient bus system that is convenient for riders. As articulated, for example, in the Downs Thomson paradox car traffic will continue to worsen until there are other comparable modes of alternative transit to get where you want to go. Therefore, if I at best can pick up a route every thirty minutes and at worst every two hours, if I have the means, I'm going to take a car to get where I need to go on time. 

```{r, warning = FALSE, message = FALSE}
# uh oh it isn't displaying properly, let's break that up so routes other 
##[2]than the one overlaid on top show.


headways_route_filter_low <- subset(headways_routes_filter, 
                                    head_cat == "low", select = c("geometry"))
headways_route_filter_low

headways_route_filter_medium <- subset(headways_routes_filter, 
                                       head_cat == "medium", select = c("geometry"))
headways_route_filter_medium

headways_route_filter_high <- subset(headways_routes_filter, 
                                     head_cat == "high", select = c("geometry"))
headways_route_filter_high
```

### Plotting Headways Map
```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}

par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(headways_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(headways_route_filter_medium %>% st_geometry(), col = "yellow2", add = TRUE)
plot(headways_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
title(main = "Routes Categorized by Frequency of Service")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "yellow2", "red"), bty = "n")
```

## Candidates for Bus Rapid Transit

While perhaps simplistic or brute this map, displays the best candidates for 
increased investment. Most modern systems have bus rapid transit (BRT) lines, which
run frequently enough that an end user shouldn't have to worry about 
timetables.Basic geostatistic analysis in my other class indicates 
upwards of 13,000 bus rider candidates who would benefit from a BRT line 
on the 27 alone with 48,000 residents in walking distance. Future
iterations of this analysis will hopeful indicate the same thing.

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}

par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(subsidy_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(distinct_route_filter_high %>% st_geometry(), col = "green", add = TRUE)
title("Routes that are Good Candidates for BRT")
```


## Candidates for Removal

The routes above could be subsidized by removing the following routes which have
extremely high subsidy, over three times the system average, and low ridership.
The top three worst routes are routes 8, 22, and 29.

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}

par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(headways_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
plot(subsidy_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
plot(distinct_route_filter_low %>% st_geometry(), col = "red", add = TRUE)
title("Candidates for Route Removal")
```

