---
title: "Garren"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Garren}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following is a vignette of Garren Kalter's work for the final project. As you can see,
the following is predominantly vector data, that has been cut to the city boundaries of Worcester, joined with relevant preprocessed data about the routes, and then displayed to highlight charateristic discrepancies between them. 

## Reading in Relevant Libraries

```{r}
library(wrta)
library(sf)
library(dplyr)
library(rgdal)
library(geospaar)
```


## Reading in Relevant Data

```{r, warning = FALSE, message = FALSE }
ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% 
  readxl::read_excel() 

county <- system.file("extdata/tl_2017_25027_faces/tl_2017_25027_faces.shp",
                     package = "wrta") %>% st_read()

roads <- system.file("extdata/tl_2018_25027_roads/tl_2018_25027_roads.shp",
                     package = "wrta") %>% st_read()

stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()

routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% st_read() %>% st_zm()

worcester <- system.file("extdata/Worcester/worcester.shp",
                     package = "wrta") %>% st_read()

```

## Cropping

```{r, warning = FALSE, message = FALSE}
#reproject
routes <- st_transform(x = routes, crs = st_crs(worcester))

stops <- st_transform(x = stops,crs = st_crs(worcester))

roads <- st_transform(x = roads, crs = st_crs(worcester))

#major roads in Worcester
roads_worc <- st_intersection(x = roads, y = worcester) 
#bus stops
stops_worc <- st_intersection(x = stops, y = worcester)
#bus routes
routes_worc <- st_intersection(x = routes, y = worcester)

```

## Join

```{r, warning = FALSE, message = FALSE}
#Identifying whether variables Match
routes$Bus_Route %in% ridership$ROUTE

#completing a join via coercion
routes_ridership <- right_join(x = ridership, 
                               y = routes_worc, by = c("ROUTE" = "BusRoute")) 
```

## Categories Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
#cleaning data up
distinct_routes <- routes_ridership %>% distinct(ROUTE, .keep_all = TRUE) 

#generating a new field to categorize ridership
distinct_routes_filter$ride_cat
distinct_routes_filter <- distinct_routes %>% 
  mutate(ride_cat = ifelse(WKDAY_RIDERS > 170000, "high", "other"), 
         ride_cat = ifelse(between(WKDAY_RIDERS, 10000, 170000), "medium", ride_cat),
         ride_cat = ifelse(WKDAY_RIDERS < 10000, "low", ride_cat)) %>% 
  st_as_sf()
```

### Correcting for Overlaid Routes

```{r, warning = FALSE, message = FALSE}

# uh oh it isn't displaying properly, let's break that up so routes other 
##[2]than the top one shows.

distinct_route_filter_low <- subset(distinct_routes_filter, 
                                    ride_cat == "low", select = c("geometry")) 

distinct_route_filter_medium <- subset(distinct_routes_filter, 
                                       ride_cat == "medium", select = c("geometry"))

distinct_route_filter_high <- subset(distinct_routes_filter, 
                                     ride_cat == "high", select = c("geometry"))
```

### Plotting Categories of Ridership

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
#plot
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(distinct_route_filter_low %>% st_geometry(), col = "red", add = TRUE)
plot(distinct_route_filter_medium %>% st_geometry(), col = "yellow", add = TRUE)
plot(distinct_route_filter_high %>% st_geometry(), col = "green", add = TRUE)
title(main = "Routes Categorized by Ridership")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("red", "yellow2", "green"), bty = "n")
```

## Subsidies Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
subsidy_routes_filter$sub_cat
subsidy_routes_filter <- distinct_routes %>% 
  mutate(sub_cat = ifelse(subsidy > 10.00, "high", "other"), 
         sub_cat = ifelse(between(subsidy, 3.00, 10.00), "medium", sub_cat),
         sub_cat = ifelse(subsidy < 3.00, "low", sub_cat)) %>% 
  st_as_sf()
```

### Correcting for Overlaid Routes

```{r, warning = FALSE, message = FALSE}
subsidy_route_filter_low <- subset(subsidy_routes_filter, 
                                   sub_cat == "low", select = c("geometry")) 

subsidy_route_filter_medium <- subset(subsidy_routes_filter, 
                                      sub_cat == "medium", select = c("geometry"))

subsidy_route_filter_high <- subset(subsidy_routes_filter, 
                                    sub_cat == "high", select = c("geometry"))
```

### Plotting Subsidies Map

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
#plot
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(subsidy_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(subsidy_route_filter_medium %>% st_geometry(), col = "yellow2", add = TRUE)
plot(subsidy_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
title(main = "Routes Categorized by Subsidy")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "yellow2", "red"), bty = "n")

```


## Headways Map Mutation

```{r, warning = FALSE, message = FALSE, error = TRUE}
headways_routes_filter$head_cat
headways_routes_filter <- distinct_routes %>% 
  mutate(head_cat = ifelse(time > 60, "high", "other"), 
         head_cat = ifelse(between(time, 30, 60), "medium", head_cat),
         head_cat = ifelse(time < 30, "low", head_cat)) %>% 
  st_as_sf()
```

### Correcting for Overlaid Routes

```{r, warning = FALSE, message = FALSE}
headways_route_filter_low <- subset(headways_routes_filter, 
                                    head_cat == "low", select = c("geometry")) 

headways_route_filter_medium <- subset(headways_routes_filter, 
                                       head_cat == "medium", select = c("geometry"))

headways_route_filter_high <- subset(headways_routes_filter, 
                                     head_cat == "high", select = c("geometry"))
```

### Plotting Headways Map
```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
# as you can see there are no routes which run below 30 minutes, this is a problem
# when trying to run an efficient WRTA
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(headways_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(headways_route_filter_medium %>% st_geometry(), col = "yellow2", add = TRUE)
plot(headways_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
title(main = "Routes Categorized by Frequency of Service")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "yellow2", "red"), bty = "n")

```

## Candidates for Bus Rapid Transit
```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
# While perhaps simplistic or brute this map, displays the best candidates for 
# [2] increased investment. Most modern bus ways have rapid transit line, which
# [3] run frequently enough that an end user shouldn't have to worry about 
# [4] timetables.Basic geostatistic analysis in my other class indicates 
# [5] upwards of 13,000 bus rider candidates who would benefit from a BRT line 
# [6] on the 27 alone with 48,000 residents in walking distance.Future
# [7] iterations of this analysis will hopeful indicate the same thing.

par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(subsidy_route_filter_low %>% st_geometry(), col = "green", add = TRUE)
plot(distinct_route_filter_high %>% st_geometry(), col = "green", add = TRUE)
title("Routes that are Good Candidates for BRT")


```


Candidates for Removal

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
# The routes above could be subsidized by removing the following routes which have
# [2] extremely high subsidy, over three times the system average, and low ridership. 
par(mar = c(0, 0, 1, 0))
plot(worcester %>% st_geometry(), col = "grey")
plot(headways_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
plot(subsidy_route_filter_high %>% st_geometry(), col = "red", add = TRUE)
title("Routes that Should be Evaluated for Closure")

```

