---
title: "Luke"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Luke}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following vignette is Luke's code for the project. Due to a Github complication,
Garren uploaded and cleaned the code. The following vignette displays basic information
regarding the population density and concentration of vehicles in Worcester, MA. This is important for our analysis as they are strong indicators of where potential bus riders
might be in the city. 

* Many outputs have been shortened using head() in order to improve readability while
still allowing the end user to see the outputs for themselves. A few have been left out
for redundancies sake and/or where doing so didn't make sense.

## Reading in Packages and Libraries

```{r setup}
library(dplyr)
library(ggplot2)
library(rgdal)
library(raster)
library(sf)
library(wrta)
```

## Reading in Files

```{r}
# Read in files
ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% 
  readxl::read_excel() 

roads <- system.file("extdata/tl_2018_25027_roads/tl_2018_25027_roads.shp",
                     package = "wrta") %>% st_read()

stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()

routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% st_read() %>% st_zm()

census_poly<- system.file("extdata/Census/export-gisdata.mapc.census2010_tracts_poly.shp",          package = "wrta") %>% st_read()

vehicle <- system.file("extdata/tabular.trans_mavc_public_summary_ct_2014_q4.csv", package = "wrta") %>% readr::read_csv(.)

census_pop <- system.file("extdata/tabular.demo_general_demographics_ct_2010.csv", package = "wrta") %>% readr::read_csv(.)

worcester <- system.file("extdata/Worcester/worcester.shp",
                     package = "wrta") %>% st_read()
```

## Reprojection and Cropping

```{r, warning = FALSE, message = FALSE }

#Reproject to same CRS
routes <- st_transform(x = routes, crs = st_crs(worcester))
head(routes)

stops <- st_transform(x = stops, crs = st_crs(worcester)) 
head(stops)

roads <- st_transform(x = roads, crs = st_crs(worcester))
roads

census_poly <- st_transform(x = census_poly, crs = st_crs(worcester)) 
head(census_poly)

# Cropping
#major roads in Worcester
roads_worc <- st_intersection(x = roads, y = worcester) 
head(roads_worc)
#bus stops
stops_worc <- st_intersection(x = stops, y = worcester) 
head(stops_worc)
#bus routes
routes_worc <- st_intersection(x = routes, y = worcester)
head (routes_worc)
```

## Joining and Cleaning Dataframes

```{r,  warning = FALSE, message = FALSE}
#join data frame by census tract id
census_data <- left_join(census_pop, vehicle, by = "ct10_id")  #
head(census_data)
## Polygon convert character to numeric column
census_data_pop <- census_poly %>% mutate(ct10_id = as.numeric(ct10_id))
head(census_data_pop)
#joining new dataset
census_data_poly <- left_join(census_data_pop, census_data, by = "ct10_id") 
head(census_data_poly)
#census worcester data 
census_worc <- st_intersection(x = census_data_poly, y = worcester)  
head(census_worc)
#clean up population data
pop_density <- census_worc[ , c("pop", "area_acres", "area_sqft", "ct10_id", "objectid", "shape")] 
pop_density #pop density data worcester

#clean up vehicle data
vehicle_density <- census_worc[ , c("veh_tot", "area_acres", "area_sqft", "ct10_id", "objectid", "shape")] 
vehicle_density # vehicle density data worcester


```

## Mutations

```{r}
# adding fields regarding density
#no output displayed

#density over sq acres
pop_density$density_acres <- (pop_density$pop/640)

# density over sq mile
pop_density$density_sqmile <- (pop_density$pop/vehicle_density$area_sqft) 

# density over sq acre
vehicle_density$density_acres <- (vehicle_density$veh_tot/640) 

#density over sq mile
vehicle_density$density_sqmile <- (vehicle_density$veh_tot/vehicle_density$area_sqft)
```

## Subsetting Data
```{r}
vehicle_density_subset <- subset(vehicle_density, select = c("density_sqmile", "geometry"))  
head(vehicle_density_subset)

pop_density_subset <- subset(pop_density, select = c("density_sqmile", "geometry"))
head(pop_density_subset)
```

## Calculating Mean
```{r}
mean(pop_density_subset$density_sqmile)

mean(vehicle_density_subset$density_sqmile)
```

## Population Density Plot

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
par(mar = c(0, 0, 1, 0)) 
plot(x = pop_density_subset %>% st_geometry(),  col = c("orange", "red", "bisque"))
title(main = "Population Density of Worcester")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("orange", "red", "bisque"), bty = "n")
```


## Vehicle Density Plot

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center", error = TRUE }
par(mar = c(0, 0, 1, 0)) 
plot(vehicle_density_subset %>% st_geometry(), col = c("green", "blue", "aquamarine"))
title(main = "Vehicle Density of Worcester")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "blue", "aquamarine"), bty = "n")
```


