---
title: "Luke"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Luke}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following vignette is Luke's code for the project. Due to a Github complication,
Garren uploaded and cleaned the code. The following vignette displays basic information
regarding the population density and concentration of vehicles in Worcester, MA. This is important for our analysis as they are strong indicators of where potential bus riders
might be in the city. 

## Reading in Packages and Libraries

```{r setup}
library(dplyr)
library(ggplot2)
library(rgdal)
library(raster)
library(sf)
library(wrta)
```

## Reading in Files

```{r}
# Read in files
ridership <- system.file("extdata/Weekday_ridership.xlsx", package = "wrta") %>% 
  readxl::read_excel() 

roads <- system.file("extdata/tl_2018_25027_roads/tl_2018_25027_roads.shp",
                     package = "wrta") %>% st_read()

stops <- system.file("extdata/WRTA_August_2020/Active_WRTA_Bus_Stops_Aug2020.shp",
                     package = "wrta") %>% st_read()

routes <- system.file("extdata/WRTA_August_2020/Active_WRTA_Routes_Aug2020.shp",
                      package = "wrta") %>% st_read() %>% st_zm()

census_poly<- system.file("extdata/Census/export-gisdata.mapc.census2010_tracts_poly.shp",          package = "wrta") %>% st_read()

vehicle <- system.file("extdata/tabular.trans_mavc_public_summary_ct_2014_q4.csv", package = "wrta") %>% readr::read_csv(.)

census_pop <- system.file("extdata/tabular.demo_general_demographics_ct_2010.csv", package = "wrta") %>% readr::read_csv(.)

worcester <- system.file("extdata/Worcester/worcester.shp",
                     package = "wrta") %>% st_read()
```

## Reprojection and Cropping

```{r}
# summary statistics and joins for setup
#Cropping
#reproject
routes <- st_transform(x = routes, crs = st_crs(worcester)) # reproject to same CRS

stops <- st_transform(x = stops, crs = st_crs(worcester)) # reproject to same CRS

roads <- st_transform(x = roads, crs = st_crs(worcester))# reproject to same CRS

census_poly <- st_transform(x = census_poly, crs = st_crs(worcester)) # reproject to same CRS
#major roads in Worcester
roads_worc <- st_intersection(x = roads, y = worcester) #roads worcester intersect
#bus stops
stops_worc <- st_intersection(x = stops, y = worcester) # stops worcester intersect
#bus routes
routes_worc <- st_intersection(x = routes, y = worcester) # routes worcester intersect
```

## Joining and Cleaning Dataframes

```{r}
#join data frame by census tract id
census_data <- left_join(census_pop, vehicle, by = "ct10_id")  #

## Polygon convert character to numeric column
census_data_pop <- census_poly %>% mutate(ct10_id = as.numeric(ct10_id))

#joining new dataset
census_data_poly <- left_join(census_data_pop, census_data, by = "ct10_id") 

#census worcester data 
census_worc <- st_intersection(x = census_data_poly, y = worcester)  

#clean up population data
pop_density <- census_worc[ , c("pop", "area_acres", "area_sqft", "ct10_id", "objectid", "shape")] 
pop_density #pop density data worcester

#clean up vehicle data
vehicle_density <- census_worc[ , c("veh_tot", "area_acres", "area_sqft", "ct10_id", "objectid", "shape")] 
vehicle_density # vehicle density data worcester


```

## Mutations

```{r}
# adding fields regarding density

#density over sq acres
pop_density$density_acres <- (pop_density$pop/640)

# density over sq mile
pop_density$density_sqmile <- (pop_density$pop/vehicle_density$area_sqft) 

# density over sq acre
vehicle_density$density_acres <- (vehicle_density$veh_tot/640) 

#density over sq mile
vehicle_density$density_sqmile <- (vehicle_density$veh_tot/vehicle_density$area_sqft)
```

## Subsetting Data
```{r}
vehicle_density_subset <- subset(vehicle_density, select = c("density_sqmile", "geometry"))  

pop_density_subset <- subset(pop_density, select = c("density_sqmile", "geometry"))
```

## Calculating Mean
```{r}
mean(pop_density_subset$density_sqmile)

mean(vehicle_density_subset$density_sqmile)
```

## Population Density Plot

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center"}
par(mar = c(0, 0, 1, 0)) 
plot(x = pop_density_subset %>% st_geometry(),  col = c("orange", "red", "bisque"))
title(main = "Population Density of Worcester")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("orange", "red", "bisque"), bty = "n")
```


## Vehicle Density Plot

```{r, fig.width=4.75, fig.height=3.75, fig.align = "center", error = TRUE }
par(mar = c(0, 0, 1, 0)) 
plot(vehicle_density_subset %>% st_geometry(), col = c("green", "blue", "aquamarine"))
title(main = "Vehicle Density of Worcester")
legend(x = "bottomright", legend = c("low", "medium", "high"), 
       fill = c("green", "blue", "aquamarine"), bty = "n")
```


